Enabling GPU Virtualization in Cloud Environments

### 2016-02-09 paper.tex
motivation removed

Our proposal goes further, not only bringing solutions for all kind of HPC applications, 
but also aiming to boost flexibility in the use of GPUs.

\section{\uppercase{Motivation}}
\label{sec:motivation}

Cloud computing and storage have gained wide appeal  
during the last years thanks to a relatively easy configuration  
process and a low IT invest. The services offered in the cloud are useful for companies and individuals 
that look for temporary access to state-of-the-art technology without spending a 
huge amount of money. In general, the systems provided by cloud 
companies match those requested by the clients thanks to its flexibility.

Two different types of cloud approaches are currently available:
on one hand, public facilities, such as AWS or 
Microsoft Azure, that allow users to chose a pre-configured system.   
On the other hand, institutions/companies 
are able to deploy their own cloud infrastructure using a software as OpenStack.

In both cases, the power of current cloud facilities relies on the use 
of VMs which  
emulate a physical node with its own 
CPUs, storage, network interconnection, GPU and OS. 
However, providing a VM equipped with a GPU implies
that the guest host must be equipped with at least one real GPU, and
once it has been assigned to an instance, no other VM can use
it. 
\textcolor{black}{Therefore, the number of VMs that can access the GPUs is 
limited by the number of local devices. 
This lack of 
flexibility is also present in cluster configurations, where 
depending on the users' necessities, 
two approaches can be identified to enable GPGPU computation.
On the one hand, the conservative option integrates many hosts with GPUs to ensure that it is likely to satisfy all requests.
On the other hand, a more daring approach employs a reduced count of GPU-enabled nodes, based on an euristhic value calculated from past experience.
}

The use of a virtualization technology, such as {rCUDA}, adds 
the desired flexibility because this software detaches the GPUs 
from the nodes where they are installed, allowing all nodes in a cluster
to use them as if they were local, as well as to share them, 
achieving a higher utilization rate and reducing the idle periods.
Thanks to the ``shareability'' introduced by {rCUDA}, all VMs in the cloud 
environment can leverage and access all GPUs installed in 
the physical nodes, removing past limitations.

Our work selects AWS as a public 
solution, and OpenStack as a private cloud environment. 
In order to achieve our goal, these cloud facilities need to be integrated with the {rCUDA} software. In addition, for OpenStack, a module for managing virtualized GPUs 
needs to be implemented.

>>>>

Our proposal goes further, not only bringing solutions for all kind of HPC applications, 
but also aiming to boost flexibility in the use of GPUs.
%
%\section{\uppercase{Motivation}}
%\label{sec:motivation}
%
%Cloud computing and storage have gained wide appeal  
%during the last years thanks to a relatively easy configuration  
%process and a low IT invest. The services offered in the cloud are useful for companies and individuals 
%that look for temporary access to state-of-the-art technology without spending a 
%huge amount of money. In general, the systems provided by cloud 
%companies match those requested by the clients thanks to its flexibility.
%
%Two different types of cloud approaches are currently available:
%on one hand, public facilities, such as AWS or 
%Microsoft Azure, that allow users to chose a pre-configured system.   
%On the other hand, institutions/companies 
%are able to deploy their own cloud infrastructure using a software as OpenStack.  
%
%In both cases, the power of current cloud facilities relies on the use 
%of VMs which  
%emulate a physical node with its own 
%CPUs, storage, network interconnection, GPU and OS. 
%However, providing a VM equipped with a GPU implies
%that the guest host must be equipped with at least one real GPU, and
%once it has been assigned to an instance, no other VM can use
%it. 
%\textcolor{black}{Therefore, the number of VMs that can access the GPUs is 
%limited by the number of local devices. 
%This lack of 
%flexibility is also present in cluster configurations, where 
%depending on the users' necessities, 
%two approaches can be identified to enable GPGPU computation.
%On the one hand, the conservative option integrates many hosts with GPUs to ensure that it is likely to satisfy all requests.
%On the other hand, a more daring approach employs a reduced count of GPU-enabled nodes, based on an euristhic value calculated from past experience.
%}
%
%The use of a virtualization technology, such as {rCUDA}, adds 
%the desired flexibility because this software detaches the GPUs 
%from the nodes where they are installed, allowing all nodes in a cluster
%to use them as if they were local, as well as to share them, 
%achieving a higher utilization rate and reducing the idle periods.
%Thanks to the ``shareability'' introduced by {rCUDA}, all VMs in the cloud 
%environment can leverage and access all GPUs installed in 
%the physical nodes, removing past limitations. 
%
%Our work selects AWS as a public 
%solution, and OpenStack as a private cloud environment. 
%In order to achieve our goal, these cloud facilities need to be integrated with the {rCUDA} software. In addition, for OpenStack, a module for managing virtualized GPUs 
%needs to be implemented.
